\textbf{Advection versus diffusion in brain intraventricular transport}

### 2025-03-28 paper/main.tex
minor revisions

Including the stabilization outlined above, the semi-discrete weak formulation of
the Stokes equations reads:
find ($\uu_h, p_h$)$\in$($\VV_h, Q_h$) such that
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu_h, \vv_h)
        + \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h)
        + \mathcal{B}(\vv_h, p_h)
        &= \mathcal{L}(\vv_h), &&\quad\forall\vv_h\in\VV_h, \\
        \mathcal{B}(\uu_h, q_h) &= 0, &&\quad\forall q_h\in Q_h.
    \end{alignat}%
    \label{eq:abstract_semi_discrete_weak_form_modelC}%
\end{subequations}%

In time, we discretize the weak formulation (\Cref{eq:abstract_semi_discrete_weak_form_modelC})
with the implicit Euler method, approximating the time derivative as
\begin{equation*}
    \pdifft{\uu} \approx \frac{\uu^{n+1} - \uu^n}{\Delta t},
\end{equation*}
where $n$ denotes the timestep index. All of the other terms
in~\Cref{eq:abstract_semi_discrete_weak_form_modelC} are evaluated at timestep $n+1$.
The method is implicit and of first order in time~\cite{butcher2016numerical}. 
The fully discrete Stokes equations were solved with the penalty parameter $\gamma=10$.

@@

\subsection{Discretization of the advection-diffusion equation}
We outline the discretization of the advection-diffusion weak formulation
for the baseline model, defined
by~\Cref{eq:adv_diff_weak_formB,eq:adv_diff_skew_form,eq:adv_diff_linear_form}.
The weak formulation is discretized with a symmetric interior-penalty method using
discontinuous Galerkin finite elements~\cite{Arnold1982AnElements}.
Exchanging the variables in~\Cref{eq:adv_diff_weak_formB} with their discrete
counterparts yields the discrete weak form: find $c_h\in W_h$ such that
\begin{equation*}
    a(c_h, \phi_h) + b(c_h, \phi_h) = L(\phi_h), \quad\forall\phi_h\in W_h.
\end{equation*}
The space $W_h$ is constructed with finite elements with
discontinuous Lagrange polynomial basis functions:
\begin{equation}
    W_h = \Big\{w_h : w_h\in L^2;  w_h\vert_K\in P_k(K); \forall K\in\Omega_h\Big\}.
\end{equation}
These elements allow for discontinuities in the discrete solution across element edges.
To ensure a stable discretization with these elements, we add the stabilization
\begin{equation}
    \begin{aligned}
        a_{\mathrm{stab}}(c_h, \phi_h)
        = &\sum_{f\in \mathcal{F}}\int_f{D\frac{\alpha}{\avg{h}}\jump{c_h}\cdot\jump{\phi_h}}\,\mathrm ds
        +\sum_{f\in \mathcal{F}}\int_f{\left({c_h \hat{u}_h}|^+ - {c_h \hat{u}_h}|^-\right)\cdot\jump{\phi_h}}\,\mathrm ds  \\
         -&\sum_{f\in \mathcal{F}}\int_f{D\avg{\nabla\phi_h}\cdot\jump{c_h}}\,\mathrm ds
         -\sum_{f\in \mathcal{F}}\int_f{D\avg{\nabla c_h}\cdot\jump{\phi_h}}\,\mathrm ds,
    \end{aligned}\label{eq:SIP_stabilization}
\end{equation}
to the discrete weak formulation, where
\begin{equation*}
    \hat{u}_h = \frac{1}{2}\left(\uu_h\cdot\nn + \vert\uu_h\cdot\nn\vert\right)
\end{equation*}
is used to upwind the velocity $\uu_h$ with respect to a
facet. The stabilization parameter $\alpha>0$ must be chosen
large enough to achieve stability. In~\Cref{eq:SIP_stabilization}, the first two terms
penalize jumps in the concentration flux across interior facets,
and are needed to preserve coercivity of the bilinear form.
The last two terms in $a_{\mathrm{stab}}$ are
added to preserve symmetry and consistency of the bilinear form, respectively.
The average and jump operators used
in~\Cref{eq:SIP_stabilization} are defined, respectively, as
\begin{equation*}
    \avg{c} = \frac{c^+ + c^-}{2} \qquad\mathrm{and}\qquad \jump{c} = c^+\nn^+ + c^-\nn^-.
\end{equation*}
The final discrete weak formulation of the advection-diffusion equation is:
find $c_h\in W_h$, such that
\begin{equation}
    a(c_h, \phi_h) + a_{\mathrm{stab}}(c_h, \phi_h) + b(c_h, \phi_h) = L(\phi_h),
    \quad\forall\phi_h\in W_h.
    \label{eq:discrete_weak_form_adv_diff}
\end{equation}

In time, we discretized~\Cref{eq:discrete_weak_form_adv_diff} with the
second-order backward difference scheme BDF2~\cite{Volker2016}. This scheme
approximates the time derivative as
\begin{equation*}
    \pdifft{c} \approx \frac{3c^{n+1} - 4c^n + c^{n-1}}{2\Delta t},
\end{equation*}
where $n$ denotes the timestep index. All of the other terms
in~\Cref{eq:discrete_weak_form_adv_diff} are evaluated at timestep $n+1$.
The method is implicit and of second order in time. We solved the fully
discrete weak formulation of the advection-diffusion equation
with the penalty parameter value $\alpha=50$.

>>>>

Including the stabilization outlined above, the semi-discrete weak formulation of
the Stokes equations reads:
find ($\uu_h, p_h$)$\in$($\VV_h, Q_h$) such that
\begin{subequations}
    \begin{alignat}{2}
        \mathcal{A}(\uu_h, \vv_h)
        + \mathcal{A}_{\mathrm{stab}}(\uu_h, \vv_h)
        + \mathcal{B}(\vv_h, p_h)
        &= \mathcal{L}(\vv_h), &&\quad\forall\vv_h\in\VV_h, \\
        \mathcal{B}(\uu_h, q_h) &= 0, &&\quad\forall q_h\in Q_h.
    \end{alignat}%
    \label{eq:abstract_semi_discrete_weak_form_modelC}%
\end{subequations}%
In time, we discretize the weak formulation (\Cref{eq:abstract_semi_discrete_weak_form_modelC})
with the implicit Euler method, approximating the time derivative as
\begin{equation*}
    \pdifft{\uu} \approx \frac{\uu^{n+1} - \uu^n}{\Delta t},
\end{equation*}
where $n$ denotes the timestep index. All of the other terms
in~\Cref{eq:abstract_semi_discrete_weak_form_modelC} are evaluated at timestep $n+1$.
The method is implicit and of first order in time~\cite{butcher2016numerical}. 
The fully discrete Stokes equations were solved with the penalty parameter $\gamma=10$.

@@

\subsection{Discretization of the advection-diffusion equation}
We outline the discretization of the advection-diffusion weak formulation
for the baseline model, defined
by~\Cref{eq:adv_diff_weak_formB,eq:adv_diff_skew_form,eq:adv_diff_linear_form}.
The weak formulation is discretized with a symmetric interior-penalty method using
discontinuous Galerkin finite elements~\cite{Arnold1982AnElements}.
Exchanging the variables in~\Cref{eq:adv_diff_weak_formB} with their discrete
counterparts yields the discrete weak form: find $c_h\in W_h$ such that
\begin{equation*}
    a(c_h, \phi_h) + b(c_h, \phi_h) = L(\phi_h), \quad\forall\phi_h\in W_h.
\end{equation*}
The space $W_h$ is constructed with finite elements with
discontinuous Lagrange polynomial basis functions:
\begin{equation}
    W_h = \Big\{w_h : w_h\in L^2;  w_h\vert_K\in P_k(K); \forall K\in\Omega_h\Big\}.
\end{equation}
These elements allow for discontinuities in the discrete solution across element edges.
To ensure a stable discretization with these elements, we add the stabilization
\begin{equation}
    \begin{aligned}
        a_{\mathrm{stab}}(c_h, \phi_h)
        = &\sum_{f\in \mathcal{F}}\int_f{D\frac{\alpha}{\avg{h}}\jump{c_h}\cdot\jump{\phi_h}}\,\mathrm ds
        +\sum_{f\in \mathcal{F}}\int_f{\left({c_h \hat{u}_h}|^+ - {c_h \hat{u}_h}|^-\right)\cdot\jump{\phi_h}}\,\mathrm ds  \\
         -&\sum_{f\in \mathcal{F}}\int_f{D\avg{\nabla\phi_h}\cdot\jump{c_h}}\,\mathrm ds
         -\sum_{f\in \mathcal{F}}\int_f{D\avg{\nabla c_h}\cdot\jump{\phi_h}}\,\mathrm ds,
    \end{aligned}\label{eq:SIP_stabilization}
\end{equation}
to the discrete weak formulation, where
\begin{equation*}
    \hat{u}_h = \frac{1}{2}\left(\uu_h\cdot\nn + \vert\uu_h\cdot\nn\vert\right)
\end{equation*}
is used to upwind the velocity $\uu_h$ with respect to a
facet. The stabilization parameter $\alpha>0$ must be chosen
large enough to achieve stability. In~\Cref{eq:SIP_stabilization}, the first two terms
penalize jumps in the concentration flux across interior facets,
and are needed to preserve coercivity of the bilinear form.
The last two terms in $a_{\mathrm{stab}}$ are
added to preserve symmetry and consistency of the bilinear form, respectively.
The average and jump operators used
in~\Cref{eq:SIP_stabilization} are defined, respectively, as
\begin{equation*}
    \avg{c} = \frac{c^+ + c^-}{2} \qquad\mathrm{and}\qquad \jump{c} = c^+\nn^+ + c^-\nn^-.
\end{equation*}

By including the stabilization $a_{\mathrm{stab}}(c_h, \phi_h)$,
the final semi-discrete weak formulation of the advection-diffusion equation is:
find $c_h\in W_h$, such that
\begin{equation}
    a(c_h, \phi_h) + a_{\mathrm{stab}}(c_h, \phi_h) + b(c_h, \phi_h) = L(\phi_h),
    \quad\forall\phi_h\in W_h.
    \label{eq:discrete_weak_form_adv_diff}
\end{equation}
In time, we discretized~\Cref{eq:discrete_weak_form_adv_diff} with the
second-order backward difference scheme BDF2~\cite{Volker2016}. This scheme
approximates the time derivative as
\begin{equation*}
    \pdifft{c} \approx \frac{3c^{n+1} - 4c^n + c^{n-1}}{2\Delta t},
\end{equation*}
where $n$ denotes the timestep index. All of the other terms
in~\Cref{eq:discrete_weak_form_adv_diff} are evaluated at timestep $n+1$.
The method is implicit and of second order in time. We solved the fully
discrete weak formulation of the advection-diffusion equation
with the penalty parameter value $\alpha=50$.
